---
author: "Jerad Acosta"
title: "Reducing Brooklyn Accidents with NYC Open Data"
date: "March 2017"
output:
  ioslides_presentation:
    css: slides.css
    logo: images/ConnectwMe.png
    self_contained: no
    incremental: true
---

# Overview / Purpose?   

## Goal

**Reducing Accidents in Brooklyn**    

The analysis for this deck and the accompanying [Data Exploration R Notebook](http://jerad.xyz/projects/flatly-float.html) and Dashboard was primarily generated from the NYC Open Data [Vehicle Collision Data Set](https://data.cityofnewyork.us/Public-Safety/NYPD-Motor-Vehicle-Collisions/h9gi-nx95).  

### The important factors influencing this work include:

- City Council Stake Holders
    + Can leverage the greatest amount of influence through their careers in local law and policy   
- Reduce Accidents in Brooklyn    
- Understand The NYC Collosion Data   

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

